---
title: "R Notebook for data cleaning"
subtitle: Project covidistress
output:
  html_notebook:
    toc: yes
    toc_float: true
---

# Load libraries
```{r}
library(pacman)

p_load(readr, tidyverse, stringr, multicon, psych)
```

# Load dataset
```{r}
d <- read_csv("covid.csv") #Relevant file name here
```

# Data cleaning
Double check why variables are read into R as character
```{r}
#Remove question info & only keep actual data
d <- d[-c(1:2), ]

#View data structure
str(d)
```

## Obtain basic demographic information
Just double-checked the survey, the response options still include blank(dash) responses. I am unsure how it's coded in the final qualtric export. Need to compare and revise the factor conversion here..
```{r}
#Change categorical variables to factors
d <- d %>% 
  mutate(Dem_gender = factor(Dem_gender, levels = c(1:3), labels = c("Male", "Female", "Other/Undisclosed")),
         Dem_edu = factor(Dem_edu, levels = c(1:7), labels = c("Doctorate", "College/Master degree", "Some college or equivalent", "Up to 12 years of school", "Up to 9 years of school", "Up to 6 years of school", "None")),
         Dem_edu_mom = factor(Dem_edu_mom, levels = c(1:7), labels = c("Doctorate", "College/Master degree", "Some college or equivalent", "Up to 12 years of school", "Up to 9 years of school", "Up to 6 years of school", "None")),
         Dem_employment = factor(Dem_employment, levels = c(1:6), labels = c("Student", "Full time employed", "Part time employed", "Self employed", "Not employed", "Retired")),
         Dem_maritalstatus = factor(Dem_maritalstatus),
         Dem_riskgroup = factor(Dem_riskgroup, levels = c(1:3), labels = c("Yes", "No", "Not sure"))) 
```

```{r}
dem <- d[, grep("Dem", names(d))]

describe(dem)
```

```{r}
names(d)
```

# Create composit scores 
## PSS10
```{r}
PSS10set <- d[, grep("PSS10", names(d))]
PSS10list <- list(PSS10_avg = c(1:3, -4, -5, 6, -7, -8, 9:13)) 
PSS10score <- scoreTest(PSS10set, PSS10list, nomiss = 0.01, rel = F)
d <- data.frame(d, PSS10score)
```

```{r}
#reliability of the scale
alpha(PSS10set, keys = c("Scale_PSS10_UCLA_4",
                         "Scale_PSS10_UCLA_5",
                         "Scale_PSS10_UCLA_7",
                         "Scale_PSS10_UCLA_8"))
```

## Compliance
```{r}
Compset <- d[, grep("Compliance", names(d))]
complist <- list(Compliance_avg = c(1:3, -4, 5, -6)) #Not sure if buying large groceries is against the recommendations, I'd say yes?
Compscore <- scoreTest(Compset, Complist, nomiss = 0.01, rel = F)
d <- data.frame(d, Compscore)
```

```{r}
#reliability of the scale
alpha(Compset, keys = c("Compliance_4",
                        "Compliance_6"))
```

## BFF_15
```{r}
BFF15set <- d[, grep("BFF_15", names(d))]
BFF15list <- list(neu = c(1, 2, -3), 
                  ext = c(4, 5, -6),
                  ope = c(7, 8, 9),
                  agr = c(10, 11, -12),
                  con = c(13, -14, 15)) 
BFF15score <- scoreTest(BFF15set, BFF15list, nomiss = 0.01, rel = F)
d <- data.frame(d, BFF15score)
```

```{r}
#reliability of the scale
alpha(BFF15set[1:3], keys = "BFF_15_3")
alpha(BFF15set[4:6], keys = "BFF_15_6")
alpha(BFF15set[7:9])
alpha(BFF15set[10:12], keys = "BFF_15_12")
alpha(BFF15set[13:15], keys = "BFF_15_14")
```

## SPS10
```{r}
SPS10set <- d[, grep("SPS", names(d))]
SPS10list <- list(SPS_avg = c(1:3, -4, 5, -6)) 
SPS10score <- scoreTest(SPS10set, SPS10list, nomiss = 0.01, rel = F)
d <- data.frame(d, SPS10score)
```

```{r}
#reliability of the scale
alpha(SPS10set)
```

