---
title: "R Notebook for data cleaning"
subtitle: Project covidistress - fourth data extraction
output:
  html_notebook:
    toc: yes
    toc_float: yes
  html_document:
    df_print: paged
    toc: yes
---



# Load libraries
```{r, message = FALSE}
library(pacman)

p_load(qualtRics, tidyverse, stringr, multicon, psych)
```

# Load dataset
```{r, message = FALSE}
# The read_survey function from qualtRics package keeps the header and removes the first two non-data rows 
# If you want to upload newer versions of the data, please download the .csv files from https://osf.io/z39us/files/ and add the valid appropriate file name. 
# This notebook was created for the choice version of the data files.
d <- read_survey("COVIDiSTRESS global survey May 11 2020 (choice text).csv")
```

# Data cleaning
## Filter out cases without consent and younger than 18
```{r}
d <- filter(d, Consent == "Yes") %>% 
  filter(Dem_age >= 18)
```


## Defining the time-zone used
```{r}
#Order by end date
d <- d[rev(order(as.POSIXct(d$StartDate))),]

#Change R time zone to UTC
Sys.setenv(TZ='UTC')
```

## Address issues with different languages
### User Language: Bulgarian (BG)

The variable `Country` was misplaced for people who participated in the survey in Bulgarian language.  
The data was recoded.

```{r}
# Before recoding the Country variable for the BG language: how many people answered the survey in BG and where are they from?
err <- d %>% 
  filter(UserLanguage == "BG", StartDate <= "2020-04-08 07:48:00") %>%
  select(Country, Dem_state, StartDate, EndDate) %>% 
  arrange(StartDate)

# Looking at the state of residence & country, we can figure out the date where country is shifted for the BG language
# Error occurs after StartDate == 2020-03-28 13:30:02 and correction happens at 2020-04-08 01:53:18
```

```{r}
# Move 1 country down for people responding in BG for responses between 03/27 and 04/08
d <- d %>% 
  mutate(Country = ifelse(StartDate < "2020-04-08 01:53:18" & StartDate > "2020-03-28 13:30:02",
                          ifelse(UserLanguage == "BG",
                            case_when(Country == "- other" ~ NA_character_, 
                                    Country == "Afghanistan" ~ "Afghanistan", 
                                    Country == "Algeria" ~ "Andorra",
                                    Country == "Angola" ~ "Antigua and Barbuda",
                                    Country == "Antigua and Barbuda" ~ "Argentina",
                                    Country == "Armenia" ~ "Australia",
                                    Country == "Australia" ~ "Austria",
                                    Country == "Belarus" ~ "Belgium",
                                    Country == "Brunei" ~ "Bulgaria",
                                    Country == "Bulgaria" ~ "Burkina Faso",
                                    
                                    Country == "Cameroon" ~ "Canada",
                                    Country == "Cuba" ~ "Cyprus",
                                    Country == "Cyprus" ~ "Czech Republic",
                                    Country == "Czech Republic" ~ "Denmark",
                                    Country == "Fiji" ~ "Finland",
                                    Country == "Finland" ~ "France",
                                    Country == "Georgia" ~ "Germany",
                                    Country == "Ghana" ~ "Greece",
                                    Country == "Iraq" ~ "Ireland",
                                    Country == "Ireland" ~ "Israel",
                                    
                                    Country == "Israel" ~ "Italy",
                                    Country == "Japan" ~ "Jordan",
                                    Country == "Korea, North" ~ "Korea, South",
                                    Country == "Korea, South" ~ "Kosovo",
                                    Country == "Liechtenstein" ~ "Lithuania",
                                    Country == "Lithuania" ~ "Luxembourg",
                                    Country == "Mali" ~ "Malta",
                                    Country == "Nepal" ~ "Netherlands",
                                    Country == "Nigeria" ~ "North Macedonia",
                                    Country == "Poland" ~ "Portugal",
                                    
                                    Country == "Portugal" ~ "Qatar",
                                    Country == "Qatar" ~ "Romania",
                                    Country == "South Africa" ~ "Spain",
                                    Country == "Spain" ~ "Sri Lanka",
                                    Country == "Suriname" ~ "Sweden",
                                    Country == "Sweden" ~ "Switzerland",
                                    Country == "Tanzania" ~ "Thailand",
                                    Country == "The Bahamas" ~ "Bahrain",
                                    Country == "Tunisia" ~ "Turkey",
                                    Country == "Uganda" ~ "Ukraine",
                                    
                                    Country == "Ukraine" ~ "United Arab Emirates",
                                    Country == "United Arab Emirates" ~ "United Kingdom",
                                    Country == "United Kingdom" ~ "United States",
                                    Country == "Zimbabwe" ~ NA_character_) #Zimbabwe and other are coded as NA here 
                          , Country), Country))

```

### User Language: Afrikaans (AFR)

The variable `Country` was misplaced for people who participated in the survey in Afrikaans language.  
Thus, the data was recoded. 

```{r}
# Before recoding the Country variable for the AFR language: how many people answered the survey in BG and where are they from?
d %>% 
  filter(UserLanguage == "AFR", StartDate <= "2020-04-07 06:48:00") %>% 
  count(Country)
```

```{r}
# Move 1 country down for people responding in AFR
d <- d %>% 
  mutate(Country = ifelse(UserLanguage == "AFR" & StartDate <= "2020-04-07 06:48:00",
                          ifelse(Country == "Somalia", "South Africa", "United States"), Country))
```

### User Language: Hebrew (HE)

The variable `Country` was translated and arranged according to the Hebrew alphabetical order. Thus, the data was recoded

```{r}
# How many people answer the survey in HE?
d %>% 
  filter(UserLanguage == "HE") %>% 
  count(Country) %>% 
  arrange(desc(n))
```

```{r}
d <- d %>% 
  mutate(Country = ifelse(UserLanguage == "HE",
                          case_when(Country == "Australia" ~ "Italy",
                                    Country == "Liberia" ~ "Israel"), Country))
```

### Remove dashes in front of the "- other" responses in Country
```{r}
d$Country <- str_remove(d$Country, "- ")
```

## Filter out the cases before the official launch date 2020-03-30

We excluded the cases before the official launch date because answers up to that point of time are test answers.
Here, we kept all answers from Denmark and Kosovo (even before the official launch date), because the survey started earlier in those countries.

```{r}
test <- d %>% 
  filter(!Country %in% c("Denmark", "Kosovo"),  StartDate <= "2020-03-30 13:00:00") 

nrow(test)
# We tried removing cases outside of Denmark & Kosovo before 1pm 03/30, but ended up with over 3048 observations, which is a huge number. Thus, we settled for 00:00 03/30 
```

```{r}
# Remove cases outside of Denmark & Kosovo before 03/30
test <- d %>% 
  filter(!Country %in% c("Denmark", "Kosovo"),  StartDate <= "2020-03-30 00:00:00")

nrow(test)

d <- d %>% 
  anti_join(test)
```

Some people put the string "Please exclude - test" in the `Expl_Distress_txt` variable. We checked whether there are still some left, but none were found.

```{r}
#Is there any testing answer that we missed?
mex <- d %>% 
  filter(!is.na(Expl_Distress_txt)) %>% 
  filter(Expl_Distress_txt == "Please exclude - test")

#No
```

## Rename and recode the isolation variable
We rename and recode the isolation variable:
```{r}
d <- rename(d, Dem_isolation = Dem_islolation)
d$Dem_isolation <- str_replace(d$Dem_isolation, "^1", "Uninformative response")
```


## Rearrange marital status for some languages

The marital status variable was mixed up, with the exception of English. The variable was recoded to correct for that problem.
```{r}
# Load function 'recode_if'(Aden-Buie & Gerke, 2018)
recode_if <- function(x, condition, ...) {
  if_else(condition, recode(x, ...), x)
}

# Fix differences in scoring between English and other languages 
d <- d %>%
  mutate(Dem_maritalstatus = 
           recode_if(Dem_maritalstatus, UserLanguage != "EN", 
                     "Single" = "Other or would rather not say",
                     "Married/cohabiting" = "Single",
                     "Divorced/widowed"= "Married/cohabiting",
                     "Other or would rather not say" = "Divorced/widowed"))
```

There were some participants who had "5" in `Dem_maritalstatus`. These responses were recoded as "Uninformative response"

```{r}
# Recode 5 as Uninformative responses
d$Dem_maritalstatus <- str_replace(d$Dem_maritalstatus, "^5", "Uninformative response")
```

## Recode education level & mom's education level
```{r}
#Remove dashes in front of the response options
d$Dem_edu <- str_remove(d$Dem_edu, "- ")
d$Dem_edu_mom <- str_remove(d$Dem_edu_mom, "- ")
```

There were some participants who had "1" in `Dem_edu`. These responses were recoded as "Uninformative response"

```{r}
#Recode 1 as Uninformative responses
d$Dem_edu <- str_replace(d$Dem_edu, "^1", "Uninformative response")
d$Dem_edu_mom <- str_replace(d$Dem_edu_mom, "^1", "Uninformative response")
```

## Recode gender variable for some languages

The variable `Dem_gender` in Spain & Mexico was inverted -- Male was recorded as Female and vice versa. Thus, the variable was recoded for SSP (Espanol - Espana) and SME (Espanol - Mexico)

```{r}
d <- d %>% 
  mutate(Dem_gender = ifelse(UserLanguage %in% c("SSP", "SME"),
                             case_when(Dem_gender == "Female" ~ "Male",
                                       Dem_gender == "Male" ~ "Female",
                                       Dem_gender == "Other/would rather not say" ~ "Other/would rather not say"),
                             Dem_gender))
```

## Convert character strings to factor in all demographic variables
```{r}
#Demographic information
d[,12:25] <- d[,12:25] %>% 
  mutate_if(is.character, as.factor)

glimpse(d[, 12:25])
```

## Recode AD_Check, AD_gain and AD_loss (shorten the response and turn the character to factor)
```{r}
d <- d %>% 
  mutate(AD_gain = factor(recode(AD_gain, 
                          "路 If Program A is adopted, 200 people will be saved." = "Program A",
                   "路 If Program B is adopted, there is 1/3 probability that 600 people will be saved, and 2/3 probability that no people will be saved" = "Program B")),
         AD_loss = factor(recode(AD_loss, 
                          "路 If Program C is adopted 400 people will die." = "Program C",
                          "路 If Program D is adopted there is 1/3 probability that nobody will die, and 2/3 probability that 600 people will die." = "Program D")),
         AD_check = factor(AD_check))
```

## Convert scale responses to numeric 

### PSS10
```{r}
d <- d %>% mutate_at(
  .vars = vars(contains("PSS10")),
  .funs = recode, 
  "Never" = 1, 
  "Almost never" = 2,
  "Sometimes" = 3, 
  "Fairly often" = 4,
  "Very often" = 5
  )
```


### Corona_concerns, Compliance, BFF, SPS, Coping, Expl_media
```{r, message = FALSE}
d <- d %>% mutate_at(
  .vars = vars(matches("Corona_concerns|Compliance|BFF|SPS|Coping_\\d|Expl_media")),
  .funs = recode, 
  "Strongly disagree" = 1, 
  "Disagree" = 2,
  "Slightly disagree" = 3, 
  "Slightly agree" = 4,
  "Agree" = 5,
  "Strongly agree" = 6
)
```

### Distress scale

**Be careful with values of 99 ("Does not apply to my current situation")!**

```{r}
d <- d %>% mutate_at(
  .vars = vars(matches("Distress_\\d")),
  .funs = recode, 
  "Strongly disagree" = 1, 
  "Disagree" = 2,
  "Slightly disagree" = 3, 
  "Slightly agree" = 4,
  "Agree" = 5,
  "Strongly agree" = 6,
  "Does not apply to my current situation" = 99
  
  )
```

### Trust in the country's measures

*Take care that this scale is coded in such a way that both 0 and 10 are defined as "not appropriate" (either too little or too much) and 5 is defined as appropriate.*

```{r}
# Recoding Trust_country measure
d <- d %>% mutate(
  Trust_countrymeasure = recode(Trust_countrymeasure,
  "Too little" = 0,
  "1" = 1,
  "2" = 2,
  "3" = 3,
  "4" = 4,
  "Appropriate" = 5,
  "6" = 6,
  "7" = 7,
  "8" = 8,
  "9" = 9,
  "Too much" = 10))
```

## Rename the PSS10_11 through PSS10_13 to Perceived loneliness

PSS10_11 to PSS10_13 are actually the Perceived loneliness items

```{r}
d <- d %>% 
  rename("Scale_Lon_1" = Scale_PSS10_UCLA_11,
         "Scale_Lon_2" = Scale_PSS10_UCLA_12,
         "Scale_Lon_3" = Scale_PSS10_UCLA_13) 
```

## Create composite scores
### PSS10
```{r}

PSS10set <- d[, grep("PSS10", names(d))]
PSS10list <- list(PSS10_avg = c(1:3, -4, -5, 6, -7, -8, 9, 10)) 
PSS10score <- scoreTest(PSS10set, PSS10list, nomiss = 0.01, rel = F)
d <- data.frame(d, PSS10score)
```

```{r}
#reliability of the scale
alpha(PSS10set, keys = c("Scale_PSS10_UCLA_4",
                         "Scale_PSS10_UCLA_5",
                         "Scale_PSS10_UCLA_7",
                         "Scale_PSS10_UCLA_8"))
```

### Perceived Loneliness
```{r}
Lonset <- d[, grep("Lon", names(d))]
Lonlist <- list(Lon_avg = c(1:3)) 
Lonscore <- scoreTest(Lonset, Lonlist, nomiss = 0.01, rel = F)
d <- data.frame(d, Lonscore)
```

```{r}
#reliability of the scale
alpha(Lonset)
```

### BFF_15
```{r}
BFF15set <- d[, grep("BFF_15", names(d))]
BFF15list <- list(neu = c(1, 2, -3), 
                  ext = c(4, 5, -6),
                  ope = c(7, 8, 9),
                  agr = c(-10, 11, 12),
                  con = c(13, -14, 15)) 
BFF15score <- scoreTest(BFF15set, BFF15list, nomiss = 0.01, rel = F)
d <- data.frame(d, BFF15score)
```

```{r}
#reliability of Neuroticism
alpha(BFF15set[1:3], keys = "BFF_15_3")
```
```{r}
#Extraversion
alpha(BFF15set[4:6], keys = "BFF_15_6")
```
```{r}
#Openness
alpha(BFF15set[7:9])
```

```{r}
#Agreeableness
alpha(BFF15set[10:12], keys = "BFF_15_10")
```
```{r}
#Conscientiousness
alpha(BFF15set[13:15], keys = "BFF_15_14")
```

### SPS10
```{r}
SPS10set <- d[, grep("SPS", names(d))]
SPS10list <- list(SPS_avg = c(1:10)) 
SPS10score <- scoreTest(SPS10set, SPS10list, nomiss = 0.01, rel = F)
d <- data.frame(d, SPS10score)
```

```{r}
#reliability of the scale
alpha(SPS10set)
```

## Select the variables for the cleaned version
```{r}
cleaned_d <- d[, c(5, 7, 10, 12:143)]
```

## Remove all new lines and ";" from the text variables for nice .csv output
```{r}
cleaned_d <- cleaned_d %>% 
  mutate(Expl_Distress_txt = str_replace_all(Expl_Distress_txt, pattern = "\\n", replacement = " "),
         Expl_coping_txt = str_replace_all(Expl_coping_txt, pattern = "\\n", replacement = " "),
         Final_open = str_replace_all(Final_open, pattern = "\\n", replacement = " "),
         Expl_Distress_txt = str_replace_all(Expl_Distress_txt, pattern = ";", replacement = "\\|"),
         Expl_coping_txt = str_replace_all(Expl_coping_txt, pattern = ";", replacement = "\\|"),
         Final_open = str_replace_all(Final_open, pattern = ";", replacement = "\\|")
         )
```



```{r}
#Remove hashtag and run the code below to generate a new dataset
write.csv(cleaned_d, "COVIDiSTRESS_May_11_clean.csv")
write_rds(d, "Cleaned_file.rds")

```




