---
title: "R Notebook for data cleaning"
subtitle: Project covidistress
output:
  html_notebook:
    toc: yes
    toc_float: true
---

# Load libraries
```{r}
library(pacman)

p_load(readr, tidyverse, stringr, multicon, psych)
```

# Load dataset
```{r}
d <- read_csv("covid.csv") #Relevant file name here
```

# Data cleaning
Double check why variables are read into R as character
```{r}
#Remove question info & only keep actual data
d <- d[-c(1:2), ]

#View data structure
str(d)
```

## Obtain basic demographic information
Just double-checked the survey, the response options still include blank(dash) responses. I am unsure how it's coded in the final qualtric export. Need to compare and revise the factor conversion here..
```{r}
#Change categorical variables to factors
d <- d %>% 
  mutate(Dem_gender = factor(Dem_gender, levels = c(1:3), labels = c("Male", "Female", "Other/Undisclosed")),
         Dem_edu = factor(Dem_edu, levels = c(1:7), labels = c("Doctorate", "College/Master degree", "Some college or equivalent", "Up to 12 years of school", "Up to 9 years of school", "Up to 6 years of school", "None")),
         Dem_edu_mom = factor(Dem_edu_mom, levels = c(1:7), labels = c("Doctorate", "College/Master degree", "Some college or equivalent", "Up to 12 years of school", "Up to 9 years of school", "Up to 6 years of school", "None")),
         Dem_employment = factor(Dem_employment, levels = c(1:6), labels = c("Student", "Full time employed", "Part time employed", "Self employed", "Not employed", "Retired")),
         Dem_maritalstatus = factor(Dem_maritalstatus),
         Dem_riskgroup = factor(Dem_riskgroup, levels = c(1:3), labels = c("Yes", "No", "Not sure")))

# Country recode
d <- d %>% 
  mutate(Country = recode, "2"="Afghanistan", "3"="Albania", "4"="Algeria", "5"="Andorra", "6"="Angola", "7"="Antigua and Barbuda", "8"="Argentina", "9"="Armenia", "10"="Australia", "11"="Austria", "12"="Azerbaijan", "13"="The Bahamas", "14"="Bahrain", "15"="Bangladesh", "16"="Barbados", "17"="Belarus", "18"="Belgium", "19"="Belize", "20"="Benin", "21"="Bhutan", "22"="Bolivia", "23"="Bosnia and Herzegovina", "24"="Botswana", "25"="Brazil", "26"="Brunei", "27"="Bulgaria", "28"="Burkina Faso", "29"="Burundi", "30"="Cabo Verde", "31"="Cambodia", "32"="Cameroon", "33"="Canada", "34"="Central African Republic", "35"="Chad", "36"="Chile", "37"="China", "38"="Colombia", "39"="Comoros", "40"="Congo, Democratic Republic of the", "41"="Congo, Republic of the", "42"="Costa Rica", "43"="Côte d'Ivoire", "44"="Croatia", "45"="Cuba", "46"="Cyprus", "47"="Czech Republic", "48"="Denmark", "49"="Djibouti", "50"="Dominica", "51"="Dominican Republic", "52"="East Timor (Timor-Leste)", "53"="Ecuador", "54"="Egypt", "55"="El Salvador", "56"="Equatorial Guinea", "57"="Eritrea", "58"="Estonia", "59"="Eswatini", "60"="Ethiopia", "61"="Fiji", "62"="Finland", "63"="France", "64"="Gabon", "65"="The Gambia", "66"="Georgia", "67"="Germany", "68"="Ghana", "69"="Greece", "70"="Grenada", "71"="Guatemala", "72"="Guinea", "73"="Guinea-Bissau", "74"="Guyana", "75"="Haiti", "76"="Honduras", "77"="Hungary", "78"="Iceland", "79"="India", "80"="Indonesia", "81"="Iran", "82"="Iraq", "83"="Ireland", "84"="Israel", "85"="Italy", "86"="Jamaica", "87"="Japan", "88"="Jordan", "89"="Kazakhstan", "90"="Kenya", "91"="Kiribati", "92"="Korea, North", "93"="Korea, South", "94"="Kosovo", "95"="Kuwait", "96"="Kyrgyzstan", "97"="Laos", "98"="Latvia", "99"="Lebanon", "100"="Lesotho", "101"="Liberia", "102"="Libya", "103"="Liechtenstein", "104"="Lithuania", "105"="Luxembourg", "106"="Madagascar", "107"="Malawi", "108"="Malaysia", "109"="Maldives", "110"="Mali", "111"="Malta", "112"="Marshall Islands", "113"="Mauritania", "114"="Mauritius", "115"="Mexico", "116"="Micronesia, Federated States of", "117"="Moldova", "118"="Monaco", "119"="Mongolia", "120"="Montenegro", "121"="Morocco", "122"="Mozambique", "123"="Myanmar (Burma)", "124"="Namibia", "125"="Nauru", "126"="Nepal", "127"="Netherlands", "128"="New Zealand", "129"="Nicaragua", "130"="Niger", "131"="Nigeria", "132"="North Macedonia", "133"="Norway", "134"="Oman", "135"="Pakistan", "136"="Palau", "137"="Panama", "138"="Papua New Guinea", "139"="Paraguay", "140"="Peru", "141"="Philippines", "142"="Poland", "143"="Portugal", "144"="Qatar", "145"="Romania", "146"="Russia", "147"="Rwanda", "148"="Saint Kitts and Nevis", "149"="Saint Lucia", "150"="Saint Vincent and the Grenadines", "151"="Samoa", "152"="San Marino", "153"="Sao Tome and Principe", "154"="Saudi Arabia", "155"="Senegal", "156"="Serbia", "157"="Seychelles", "158"="Sierra Leone", "159"="Singapore", "160"="Slovakia", "161"="Slovenia", "162"="Solomon Islands", "163"="Somalia", "164"="South Africa", "165"="Spain", "166"="Sri Lanka", "167"="Sudan", "168"="Sudan, South", "169"="Suriname", "170"="Sweden", "171"="Switzerland", "172"="Syria", "173"="Taiwan", "174"="Tajikistan", "175"="Tanzania", "176"="Thailand", "177"="Togo", "178"="Tonga", "179"="Trinidad and Tobago", "180"="Tunisia", "181"="Turkey", "182"="Turkmenistan", "183"="Tuvalu", "184"="Uganda", "185"="Ukraine", "186"="United Arab Emirates", "187"="United Kingdom", "188"="United States", "189"="Uruguay", "190"="Uzbekistan", "191"="Vanuatu", "192"="Vatican City", "193"="Venezuela", "194"="Vietnam", "195"="Yemen", "196"="Zambia", "197"="Zimbabwe", "199"="other" )


```

```{r}
dem <- d[, grep("Dem", names(d))]

describe(dem)
```

```{r}
names(d)
```

# Create composit scores 
## PSS10
```{r}
PSS10set <- d[, grep("PSS10", names(d))]
PSS10list <- list(PSS10_avg = c(1:3, -4, -5, 6, -7, -8, 9:13)) 
PSS10score <- scoreTest(PSS10set, PSS10list, nomiss = 0.01, rel = F)
d <- data.frame(d, PSS10score)
```

```{r}
#reliability of the scale
alpha(PSS10set, keys = c("Scale_PSS10_UCLA_4",
                         "Scale_PSS10_UCLA_5",
                         "Scale_PSS10_UCLA_7",
                         "Scale_PSS10_UCLA_8"))
```

## Compliance
```{r}
Compset <- d[, grep("Compliance", names(d))]
complist <- list(Compliance_avg = c(1:3, -4, 5, -6)) #Not sure if buying large groceries is against the recommendations, I'd say yes?
Compscore <- scoreTest(Compset, Complist, nomiss = 0.01, rel = F)
d <- data.frame(d, Compscore)
```

```{r}
#reliability of the scale
alpha(Compset, keys = c("Compliance_4",
                        "Compliance_6"))
```

## BFF_15
```{r}
BFF15set <- d[, grep("BFF_15", names(d))]
BFF15list <- list(neu = c(1, 2, -3), 
                  ext = c(4, 5, -6),
                  ope = c(7, 8, 9),
                  agr = c(10, 11, -12),
                  con = c(13, -14, 15)) 
BFF15score <- scoreTest(BFF15set, BFF15list, nomiss = 0.01, rel = F)
d <- data.frame(d, BFF15score)
```

```{r}
#reliability of the scale
alpha(BFF15set[1:3], keys = "BFF_15_3")
alpha(BFF15set[4:6], keys = "BFF_15_6")
alpha(BFF15set[7:9])
alpha(BFF15set[10:12], keys = "BFF_15_12")
alpha(BFF15set[13:15], keys = "BFF_15_14")
```

## SPS10
```{r}
SPS10set <- d[, grep("SPS", names(d))]
SPS10list <- list(SPS_avg = c(1:3, -4, 5, -6)) 
SPS10score <- scoreTest(SPS10set, SPS10list, nomiss = 0.01, rel = F)
d <- data.frame(d, SPS10score)
```

```{r}
#reliability of the scale
alpha(SPS10set)
```

