---
title: "Technical_validation"
author: "Dominik-Borna Ćepulić"
date: '10 05 2020 '
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(knitr)
library(kableExtra)
library(psych)
library(RColorBrewer)

d <- read_rds("Cleaned_file.rds")
```


# Descriptive tables

Descriptive table should include: 
 - countries in rows (pick only countries with more than 200 participants)
 - number of participants
 - percentage of participants with more than 50% of complete answers
 - percentage of participants with more than 90% of complete answers
 - sex (percentage of women)
 - mean age of participants and SD


```{r}
# Select countries that have at least 200 participants

table_countries <- d %>% count(Country) %>% 
  filter(n>=200, !(Country %in% c(NA, "other"))) %>% arrange(-n) %>% pull(Country) # Exclude other and NAs

dt <- d %>% filter(Country %in% table_countries)
```

## General descriptive table
Create the table

```{r}
# Calculate number of NAs per participant
# Relevant variables for missing values are Dem_age:Final_open, with exception of AD_gain or AD_loss (everybody saw just one of these)
length(select(dt, Dem_age:Final_open)) - 1

# All in all we have 123 variables

# Caluculate the percentage of missing values for each participant
dt <- dt %>% mutate(Perc_full = 100 - rowSums(is.na(.))/123 * 100)

no_part <- dt %>% 
  count(Country)

sex_tab <- dt %>% 
  mutate(Dem_gender = recode(Dem_gender, "Other/would rather not say" = "Other")) %>% 
  group_by(Country, Dem_gender) %>% 
  count() %>% 
  pivot_wider(names_from = Dem_gender, values_from = n, values_fill = list(n = 0)) %>% 
  rename(Gender_NA = `NA`) %>% 
  left_join(no_part) %>% 
  mutate(Prop_female = Female / n,
         Prop_male = Male / n, 
         `Prop_gender_other/not_say` = Other /n,
         Prop_gender_NA = Gender_NA / n) %>% 
  select(Country, Prop_female:Prop_gender_NA)

edu_tab <- dt %>% 
  mutate(Dem_edu = fct_recode(Dem_edu,
                                 "College degree" = "College degree, bachelor, master",
                                 "Some college" = "Some College, short continuing education or equivalent",
                                 "PhD" = "PhD/Doctorate",
                                 "< 12 years" = "Up to 12 years of school",
                                 "< 9 years" = "Up to 9 years of school",
                                 "< 6 years" = "Up to 6 years of school",
                                 "Uninformative" = "Uninformative response")
         ) %>% 
  mutate(Dem_edu = fct_relevel(Dem_edu, c("None", "< 6 years", "< 9 years", "< 12 years", "Some college", "College degree", "PhD", "Uninformative"))) %>% 
  group_by(Country, Dem_edu) %>% 
  count() %>% 
  pivot_wider(names_from = Dem_edu, values_from = n, values_fill = list(n = 0)) %>% 
  left_join(no_part) %>% 
  mutate(Prop_none = None / n,
         Prop_6years = `< 6 years` / n, 
         Prop_9years = `< 9 years` / n, 
         Prop_12years = `< 12 years` / n, 
         Prop_some_college = `Some college` / n, 
         Prop_college = `College degree` / n, 
         Prop_PhD = PhD / n, 
         Prop_edu_NA = `NA` / n,
         Prop_uninf = Uninformative / n) %>% 
  select(Country, Prop_none:Prop_uninf)

mar_tab <- dt %>% 
   mutate(Dem_maritalstatus = recode(Dem_maritalstatus, "Other or would rather not say" = "Other/wont_say")) %>% 
  group_by(Country, Dem_maritalstatus) %>% 
  count() %>% 
  pivot_wider(names_from = Dem_maritalstatus, values_from = n, values_fill = list(n = 0)) %>% 
  rename(Marital_NA = `NA`) %>% 
  left_join(no_part) %>% 
  mutate(`Prop_divorced/widowed` = `Divorced/widowed` / n,
         `Prop_married/cohabiting` = `Married/cohabiting` / n, 
         `Prop_marital_other/not_say` = `Other/wont_say` /n,
         Prop_marital_NA = Marital_NA / n,
         Prop_single = Single / n) %>% 
  select(Country, Prop_single, `Prop_married/cohabiting`, `Prop_divorced/widowed`, `Prop_marital_other/not_say`, Prop_marital_NA)



tab_desc <- dt %>% 
  group_by(Country) %>% 
  summarize(N = n(),
            N_50 = sum(Perc_full > 50)/N,
            N_90 = sum(Perc_full > 90)/N,
            M_age = mean(Dem_age, na.rm = T),
            SD_age = sd(Dem_age, na.rm = T)
  ) %>% arrange(-N) %>% 
  ungroup()

# Join tab with sex

tab_desc <- tab_desc %>% left_join(sex_tab) %>% left_join(edu_tab) %>% left_join(mar_tab, by = "Country")


# Save tab_desc to 
writexl::write_xlsx(tab_desc, path = "Technical validation/Descriptive_table.xlsx")
```

Prepare the table for pdf

```{r}
# Rename the columns
tab_desc %>% rename(Prop_50 = N_50, 
               Prop_90 = N_90,
               `Age mean` = M_age,
               `Age SD` = SD_age) %>% 
kable(digits = 2, booktabs = T, format = "latex") %>% 
  footnote(general = "Prop_50 = Proportion of participants who have more than 50% of data; Prop_90 = Proportion of participants who have more than 90% of data",threeparttable = T) %>% save_kable(file = "Tablica.pdf" )
```


## Table for each scale

For these analyses we use the cleaned dataset.  
We calculate the reliability for:  
- PSS10  
- Loneliness scale  
- SPS10  
- BFI15


Maybe ADD NUMBER OF ITEMS PER SCALE

Calculating the alpha reliabilites for all countries 

```{r}
PSS_alpha <- function(name){
  x <- dt %>% filter(Country == name) %>% select(matches("Scale_PSS10"))
  rel_stat <- alpha(x, keys = c("Scale_PSS10_UCLA_4",
                         "Scale_PSS10_UCLA_5",
                         "Scale_PSS10_UCLA_7",
                         "Scale_PSS10_UCLA_8")
  )
  rel_stat$total$raw_alpha
}

Lon_alpha <- function(name){
  x <- dt %>% filter(Country == name) %>% select(matches("Scale_Lon"))
  rel_stat <- alpha(x)
  
  rel_stat$total$raw_alpha
}

Neu_alpha <- function(name){
   x <- dt %>% filter(Country == name) %>% select(BFF_15_1:BFF_15_3)
  rel_stat <- alpha(x, keys = "BFF_15_3")
  
  rel_stat$total$raw_alpha
}
Ext_alpha<- function(name){
   x <- dt %>% filter(Country == name) %>% select(BFF_15_4:BFF_15_6)
  rel_stat <- alpha(x, keys = "BFF_15_6")
  
  rel_stat$total$raw_alpha
}
Ope_alpha<- function(name){
   x <- dt %>% filter(Country == name) %>% select(BFF_15_7:BFF_15_9)
  rel_stat <- alpha(x)
  
  rel_stat$total$raw_alpha
}
Agr_alpha<- function(name){
   x <- dt %>% filter(Country == name) %>% select(BFF_15_10:BFF_15_12)
  rel_stat <- alpha(x, keys = "BFF_15_10")
  
  rel_stat$total$raw_alpha
}
Con_alpha <- function(name){
   x <- dt %>% filter(Country == name) %>% select(BFF_15_13:BFF_15_15)
  rel_stat <- alpha(x, keys = "BFF_15_14")
  
  rel_stat$total$raw_alpha
}

SPS_alpha <- function(name){
   x <- dt %>% filter(Country == name) %>% select(matches("SPS_\\d"))
  rel_stat <- alpha(x)
  
  rel_stat$total$raw_alpha
}

```


### Table SPS

```{r}
# How many participants have all values in SPS items?
au_sps <- dt %>% 
  filter_at(.vars = vars(matches("SPS_\\d")),
            .vars_predicate = all_vars(!is.na(.))
            ) %>% 
  count(Country)

tab_sps <- dt %>% group_by(Country) %>% 
  summarize(N = n(),
            Mean = mean(SPS_avg, na.rm = T),
            SD = sd(SPS_avg, na.rm = T),
            Min = min(SPS_avg, na.rm = T),
            Max = max(SPS_avg, na.rm = T)
  ) %>% 
  ungroup() %>% 
  mutate(Alpha = map_dbl(Country, SPS_alpha)) %>% 
  left_join(au_sps, by = "Country") %>% 
  mutate(Prop_nonmis = n/N) %>% 
  select(Country, N, Prop_nonmis, Mean:Alpha)

# Save to file
writexl::write_xlsx(tab_sps, path = "Technical validation/SPS.xlsx")


# Print to pdf
tab_sps %>% 
  kable(format = "latex", booktabs = T, digits = 2, caption = "SPS scale") %>% 
  save_kable("SPS.pdf")
```

### Table PSS

```{r}
# How many participants have all values in SPS items?
au_pss <- dt %>% 
  filter_at(.vars = vars(matches("Scale_PSS10")),
            .vars_predicate = all_vars(!is.na(.))
            ) %>% 
  count(Country)

tab_pss <- dt %>% group_by(Country) %>% 
  summarize(N = n(),
            Mean = mean(PSS10_avg, na.rm = T),
            SD = sd(PSS10_avg, na.rm = T),
            Min = min(PSS10_avg, na.rm = T),
            Max = max(PSS10_avg, na.rm = T)
  ) %>% 
  ungroup() %>% 
  mutate(Alpha = map_dbl(Country, PSS_alpha)) %>% 
  left_join(au_pss, by = "Country") %>% 
  mutate(Prop_nonmis = n/N) %>% 
  select(Country, N, Prop_nonmis, Mean:Alpha)

# Save to file
writexl::write_xlsx(tab_pss, path = "Technical validation/PSS.xlsx")


# Print to pdf
tab_pss %>% 
  kable(format = "latex", booktabs = T, digits = 2, caption = "PSS scale") %>% 
  save_kable("PSS.pdf")
```

### Table Lon

```{r}
# How many participants have all values in SPS items?
au_lon <- dt %>% 
  filter_at(.vars = vars(matches("Scale_Lon")),
            .vars_predicate = all_vars(!is.na(.))
            ) %>% 
  count(Country)

tab_lon <- dt %>% group_by(Country) %>% 
  summarize(N = n(),
            Mean = mean(Lon_avg, na.rm = T),
            SD = sd(Lon_avg, na.rm = T),
            Min = min(Lon_avg, na.rm = T),
            Max = max(Lon_avg, na.rm = T)
  ) %>% 
  ungroup() %>% 
  mutate(Alpha = map_dbl(Country, Lon_alpha)) %>% 
  left_join(au_lon, by = "Country") %>% 
  mutate(Prop_nonmis = n/N) %>% 
  select(Country, N, Prop_nonmis, Mean:Alpha)

# Save to file
writexl::write_xlsx(tab_lon, path = "Technical validation/Lon.xlsx")


# Print to pdf
tab_lon %>% 
  kable(format = "latex", booktabs = T, digits = 2, caption = "Lon scale") %>% 
  save_kable("Lon.pdf")
```

### Table Neu

```{r}
# How many participants have all values in SPS items?
au_neu <- dt %>% 
  filter_at(.vars = vars(BFF_15_1:BFF_15_3),
            .vars_predicate = all_vars(!is.na(.))
            ) %>% 
  count(Country)

tab_neu <- dt %>% group_by(Country) %>% 
  summarize(N = n(),
            Mean = mean(neu, na.rm = T),
            SD = sd(neu, na.rm = T),
            Min = min(neu, na.rm = T),
            Max = max(neu, na.rm = T)
  ) %>% 
  ungroup() %>% 
  mutate(Alpha = map_dbl(Country, Neu_alpha)) %>% 
  left_join(au_neu, by = "Country") %>% 
  mutate(Prop_nonmis = n/N) %>% 
  select(Country, N, Prop_nonmis, Mean:Alpha)

# Save to file
writexl::write_xlsx(tab_neu, path = "Technical validation/Neu.xlsx")


# Print to pdf
tab_neu %>% 
  kable(format = "latex", booktabs = T, digits = 2, caption = "Neu scale") %>% 
  save_kable("Neu.pdf")
```



### Table Ext

```{r}
# How many participants have all values in SPS items?
au_ext <- dt %>% 
  filter_at(.vars = vars(BFF_15_4:BFF_15_6),
            .vars_predicate = all_vars(!is.na(.))
            ) %>% 
  count(Country)

tab_ext <- dt %>% group_by(Country) %>% 
  summarize(N = n(),
            Mean = mean(ext, na.rm = T),
            SD = sd(ext, na.rm = T),
            Min = min(ext, na.rm = T),
            Max = max(ext, na.rm = T)
  ) %>% 
  ungroup() %>% 
  mutate(Alpha = map_dbl(Country, Ext_alpha)) %>% 
  left_join(au_ext, by = "Country") %>% 
  mutate(Prop_nonmis = n/N) %>% 
  select(Country, N, Prop_nonmis, Mean:Alpha)

# Save to file
writexl::write_xlsx(tab_ext, path = "Technical validation/Ext.xlsx")


# Print to pdf
tab_ext %>% 
  kable(format = "latex", booktabs = T, digits = 2, caption = "Ext scale") %>% 
  save_kable("Ext.pdf")
```



### Table Ope

```{r}
# How many participants have all values in SPS items?
au_ope <- dt %>% 
  filter_at(.vars = vars(BFF_15_7:BFF_15_9),
            .vars_predicate = all_vars(!is.na(.))
            ) %>% 
  count(Country)

tab_ope <- dt %>% group_by(Country) %>% 
  summarize(N = n(),
            Mean = mean(ope, na.rm = T),
            SD = sd(ope, na.rm = T),
            Min = min(ope, na.rm = T),
            Max = max(ope, na.rm = T)
  ) %>% 
  ungroup() %>% 
  mutate(Alpha = map_dbl(Country, Ope_alpha)) %>% 
  left_join(au_ope, by = "Country") %>% 
  mutate(Prop_nonmis = n/N) %>% 
  select(Country, N, Prop_nonmis, Mean:Alpha)

# Save to file
writexl::write_xlsx(tab_ope, path = "Technical validation/Ope.xlsx")


# Print to pdf
tab_ope %>% 
  kable(format = "latex", booktabs = T, digits = 2, caption = "Ope scale") %>% 
  save_kable("Ope.pdf")
```


### Table Agr

```{r}
# How many participants have all values in SPS items?
au_agr <- dt %>% 
  filter_at(.vars = vars(BFF_15_10:BFF_15_12),
            .vars_predicate = all_vars(!is.na(.))
            ) %>% 
  count(Country)

tab_agr <- dt %>% group_by(Country) %>% 
  summarize(N = n(),
            Mean = mean(agr, na.rm = T),
            SD = sd(agr, na.rm = T),
            Min = min(agr, na.rm = T),
            Max = max(agr, na.rm = T)
  ) %>% 
  ungroup() %>% 
  mutate(Alpha = map_dbl(Country, Agr_alpha)) %>% 
  left_join(au_agr, by = "Country") %>% 
  mutate(Prop_nonmis = n/N) %>% 
  select(Country, N, Prop_nonmis, Mean:Alpha)


# Save to file
writexl::write_xlsx(tab_agr, path = "Technical validation/Agr.xlsx")

# Print to pdf
tab_agr %>% 
  kable(format = "latex", booktabs = T, digits = 2, caption = "Agr scale") %>% 
  save_kable("Agr.pdf")
```


### Table Con

```{r}
# How many participants have all values in SPS items?
au_con <- dt %>% 
  filter_at(.vars = vars(BFF_15_13:BFF_15_15),
            .vars_predicate = all_vars(!is.na(.))
            ) %>% 
  count(Country)

tab_con <- dt %>% group_by(Country) %>% 
  summarize(N = n(),
            Mean = mean(con, na.rm = T),
            SD = sd(con, na.rm = T),
            Min = min(con, na.rm = T),
            Max = max(con, na.rm = T)
  ) %>% 
  ungroup() %>% 
  mutate(Alpha = map_dbl(Country, Con_alpha)) %>% 
  left_join(au_con, by = "Country") %>% 
  mutate(Prop_nonmis = n/N) %>% 
  select(Country, N, Prop_nonmis, Mean:Alpha)

# Save to file
writexl::write_xlsx(tab_con, path = "Technical validation/Cons.xlsx")



# Print to pdf
tab_con %>% 
  kable(format = "latex", booktabs = T, digits = 2, caption = "Con scale") %>% 
  save_kable("Cons.pdf")
```



# Visualizations

## Number participants

```{r}
pl_desc <- dt %>% 
  group_by(Country) %>% 
  summarize(N = n(),
            N_50 = sum(Perc_full > 50),
            Nx = N - N_50
  ) %>% 
  ungroup() %>% 
  top_n(n = 10, wt = N) %>% 
  mutate(Country = fct_reorder(Country, N)) %>% 
  select(-N) %>% 
  pivot_longer(cols = N_50:Nx, names_to = "Var", values_to = "Part")

country_levels <- levels(pl_desc$Country)

 
pl_desc %>% ggplot(aes(x = Country, y = Part, fill = fct_rev(Var))) + geom_bar(stat = "identity")+
  theme_bw()+
  scale_fill_manual(values = c("#31233C", "#29C7B8")) +
  ylab("Number of participants") + 
  theme(legend.title = element_blank(),
        legend.position = "none")+
  labs(title = "The number of participants in countries with largest samples",
       subtitle = "The green part of the bars represents participants who have more than 50% of data")
```


## Sex


```{r}
pl_sex <- dt %>% 
  group_by(Country, Dem_gender) %>% 
  summarize(N = n()) %>% 
  ungroup() %>% 
  filter(Country %in% table_countries[1:10]) %>% 
  mutate(Country = fct_relevel(Country, country_levels[1:10]),
         Dem_gender = fct_reorder(fct_explicit_na(Dem_gender), N, )) 

 
pl_sex %>% filter(Dem_gender != "(Missing)") %>% 
  ggplot(aes(x = Country, y = N, fill = Dem_gender)) + geom_bar(stat = "identity")+
  theme_bw()+
  scale_fill_manual(values = c( "darkred", "#31233C", "#29C7B8")) +
  ylab("Number of participants") + 
  theme(legend.title = element_blank()) +
  labs(title = "The distribution of gender in countries with largest samples")
```



## Education


```{r}
pl_edu <- dt %>% 
  mutate(Dem_edu = fct_collapse(Dem_edu,
                                `Elementary` = c("Up to 9 years of school", "Up to 6 years of school", "None"),
                                `Higher or some college` = c("Up to 12 years of school", "Some College, short continuing education or equivalent"),
                                `College or PhD` = c("College degree, bachelor, master", "PhD/Doctorate"),
                                Uninformative = c(NA, "Uninformative response"))) %>% 
  group_by(Country, Dem_edu) %>% 
  summarize(N = n()) %>% 
  ungroup() %>% 
  filter(Country %in% table_countries[1:10]) %>% 
  mutate(Country = fct_relevel(Country, country_levels[1:10]),
         Dem_edu = fct_relevel(fct_explicit_na(Dem_edu), c("Elementary", "Higher or some college", "College or PhD", "Uninformative", NA))) 

 
pl_edu %>% filter(!(Dem_edu %in% c("(Missing)", "Uninformative"))) %>% 
  ggplot(aes(x = Country, y = N, fill = Dem_edu)) + geom_bar(stat = "identity")+
  theme_bw()+
  scale_fill_manual(values = c( "darkred", "#31233C", "#29C7B8")) +
  ylab("Number of participants") + 
  theme(legend.title = element_blank()) +
  labs(title = "The distribution of education in countries with largest samples")
```




We calculate the reliability only for countries that have at least 200 participants.


```{r}
# Select all participants that have no missing values in the given variables
#partid <- d %>% select(ResponseId, matches("Scale_PSS10|Scale_Lon|SPS_\\d|BFF_15_\\d")) %>% 
#  mutate(Mis = rowSums(is.na(.))) %>% filter(Mis == 0) %>% pull(ResponseId)
#
## Count all countries that have more than 200 of such participants
#countries <- d %>% filter(ResponseId %in% partid) %>% count(Country) %>% arrange(-n) %>% filter(n >= 200) %>% pull(Country)
#
#
## Create dataframe with those countries
#
##da <- d %>% filter(Country %in% countries)

da <- d %>% filter(Country %in% table_countries[1:10])

```


```{r}
PSS_alpha <- function(name){
  x <- da %>% filter(Country == name) %>% select(matches("Scale_PSS10"))
  rel_stat <- alpha(x, keys = c("Scale_PSS10_UCLA_4",
                         "Scale_PSS10_UCLA_5",
                         "Scale_PSS10_UCLA_7",
                         "Scale_PSS10_UCLA_8")
  )
  rel_stat$total$raw_alpha
}

Lon_alpha <- function(name){
  x <- da %>% filter(Country == name) %>% select(matches("Scale_Lon"))
  rel_stat <- alpha(x)
  
  rel_stat$total$raw_alpha
}

Neu_alpha <- function(name){
   x <- da %>% filter(Country == name) %>% select(BFF_15_1:BFF_15_3)
  rel_stat <- alpha(x, keys = "BFF_15_3")
  
  rel_stat$total$raw_alpha
}
Ext_alpha<- function(name){
   x <- da %>% filter(Country == name) %>% select(BFF_15_4:BFF_15_6)
  rel_stat <- alpha(x, keys = "BFF_15_6")
  
  rel_stat$total$raw_alpha
}
Ope_alpha<- function(name){
   x <- da %>% filter(Country == name) %>% select(BFF_15_7:BFF_15_9)
  rel_stat <- alpha(x)
  
  rel_stat$total$raw_alpha
}
Agr_alpha<- function(name){
   x <- da %>% filter(Country == name) %>% select(BFF_15_10:BFF_15_12)
  rel_stat <- alpha(x, keys = "BFF_15_10")
  
  rel_stat$total$raw_alpha
}
Con_alpha <- function(name){
   x <- da %>% filter(Country == name) %>% select(BFF_15_13:BFF_15_15)
  rel_stat <- alpha(x, keys = "BFF_15_14")
  
  rel_stat$total$raw_alpha
}

SPS_alpha <- function(name){
   x <- da %>% filter(Country == name) %>% select(matches("SPS_\\d"))
  rel_stat <- alpha(x)
  
  rel_stat$total$raw_alpha
}


dat <- tibble(Country = table_countries[1:10], 
              PSS10 = map_dbl(table_countries[1:10], PSS_alpha),
              Lon = map_dbl(table_countries[1:10], Lon_alpha),
              Neu = map_dbl(table_countries[1:10], Neu_alpha),
              Ext = map_dbl(table_countries[1:10], Ext_alpha),
              Ope = map_dbl(table_countries[1:10], Ope_alpha),
              Agr = map_dbl(table_countries[1:10], Agr_alpha),
              Con = map_dbl(table_countries[1:10], Con_alpha),
              SPS = map_dbl(table_countries[1:10], SPS_alpha))


dat %>% print(n=100)
```

## Reliabilities
Visualizing the alpha coefficients in ggplot

```{r}
dat %>% pivot_longer(cols = -Country, names_to = "Scale", values_to = "alpha") %>% 
    ggplot(aes(x = fct_reorder(Scale, alpha), y = alpha, fill = fct_reorder(Scale, alpha))) + geom_bar(stat = "identity") +
  facet_wrap(~Country, nrow = 2)+
  xlab("Scale")+
  ylab("Cronbach's alpha")+
  scale_y_continuous(breaks = seq(0, 0.9, 0.2))+
  #scale_fill_manual(values = brewer.pal(8,"BrBG"))+
  scale_fill_manual(values = c('#29c7b8', '#37aea5', '#3d9592', '#407d80', '#3f656e', '#3d4f5d', '#38394c', '#31233c'))+
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
        legend.position = "none")

```

